---
Author: Ginny Gao
Date: Feb 03, 2018
Output:
  pdf_document: default
  html_document: default
Title: SPOOKY Data Analysis
output:
  html_document: default
  pdf_document: default
---

# Goal

Goal of this project is to find characteristics of texts from 3 popular horror authors, identify similarities and differences in their texts in the [spooky](https://github.com/GU4243-ADS/spring2018-project1-ginnyqg/blob/master/data/spooky.csv) dataset. Data consists of excerpts of texts written by [Edgar Allan Poe (EAP)](https://en.wikipedia.org/wiki/Edgar_Allan_Poe), [HP Lovecraft (HPL)](https://en.wikipedia.org/wiki/H._P._Lovecraft), and [Mary Wollstonecraft Shelley (MWS)](https://en.wikipedia.org/wiki/Mary_Shelley).


# Load packages and read the data

## Setup the libraries if not already installed


```{r, message = F, warning = F}
packages.used <- c("ggplot2", "plotrix", "waffle", "dplyr", "tibble", "tidyr",  "stringr", "tidytext", "topicmodels", "wordcloud", "plotly", "webshot", "htmlwidgets", "reshape2")

# check packages that need to be installed.
packages.needed <- setdiff(packages.used, intersect(installed.packages()[,1], packages.used))

# install additional packages
if(length(packages.needed) > 0) {
  install.packages(packages.needed, dependencies = TRUE, repos = 'http://cran.us.r-project.org')
}


library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
library(stringr)
library(tidytext)
library(topicmodels)
library(wordcloud)
library(plotrix)
library(waffle)
library(plotly)
library(webshot)
library(htmlwidgets)
library(reshape2)
```


## Read in the data
`spooky.csv` in `data` folder, and this Rmd inside `doc` folder.

```{r}
spooky <- read.csv('../data/spooky.csv', as.is = TRUE)
```


# Overview of the dataset

Take a look of first few rows and dimension of the dataset

```{r}
head(spooky, 3)
dim(spooky)
```

How many texts do each author have in the dataset?

```{r}
num_texts <- table(spooky$author)
num_texts
```

Plot composition of number of texts from 3 authors in pie chart, display counts and percentages


```{r}
lbls <- paste(names(num_texts), '\n', num_texts, '\n', round(num_texts/sum(num_texts) * 100, 1), '%', sep = '')
pie3D(num_texts, labels = lbls, explode = 0.05, labelcex = 0.8)
```

# Writing Style
## Do some authors use more questions in the texts than others?


* Count number of question marks in texts for spooky  
* Add a field `num_qns` for the counts  
* Wrangle data to show counts for each author  
* Plot a waffle chart to see comparison of use of questions in texts among 3 authors.


```{r}
str_count(spooky, '\\?')
dat1 <- mutate(spooky, num_qns = str_count(spooky$text, '\\?'))
dat2 <- aggregate(dat1$num_qns, by = list(Author = dat1$author), FUN = sum)
dat2

waffle(c('EAP' = dat2[1, 2], 'HPL' = dat2[2, 2], 'MWS' = dat2[3, 2]), rows = 20, size = 0.5, title = 'Count of Questions in Texts by Authors', xlab = '(1 square == 1 question)')
```

# Sentiment analysis
## Positive and negative emotional content comparison in authors' text

Apply sentiment analysis using [bing lexicon](https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html)


```{r}
get_sentiments("bing")

tidy_text <- unnest_tokens(spooky, word, text)
tidy_text_sentiment <- tidy_text %>% inner_join(get_sentiments('bing'))
head(tidy_text_sentiment, 10)


dat3 <- table(tidy_text_sentiment$sentiment, tidy_text_sentiment$author)
dat3


pyramid.plot(dat3[1,c(1:3)], dat3[2,c(1:3)], top.labels = NULL, show.values = TRUE, ndig = 0, main = 'Author by Sentiments', unit = c('Negative', 'Positive'), ppmar = c(4, 4, 4, 4), laxlab = FALSE, raxlab = FALSE)
legend('topright', legend = c("EAP", "HPL", "MWS"), col = c("red", "green", "blue"), lty = 1, bty = 'n', lwd = 8, cex = 0.8, horiz=TRUE)
```

What are the top 100 positive and negative words do the authors use?

```{r}
#comparison cloud
tidy_text %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("darkgreen", "purple"),
                   max.words = 100)
```


* Aggregate number of negative and positive words used by each author
* Plot a bubble chart to show sentiments and num_qns by authors

```{r}
dat4 <- as.data.frame.matrix(table(tidy_text_sentiment$author, tidy_text_sentiment$sentiment))
dat4 <- setNames(cbind(rownames(dat4), dat4, row.names = NULL), c('Author', 'negative', 'positive'))
dat4

#rename column name
colnames(dat2)[which(names(dat2) == 'x')] <- 'num_qns'
dat2

#join tables by author
dat5 <- inner_join(dat2, dat4, by = 'Author')
dat5

#   Author num_qns negative positive
# 1    EAP     510     7203     6144
# 2    HPL     169     7605     3731
# 3    MWS     419     8150     6799

#plot bubble chart: Sentiments and Num of Questions per Author
p <- plot_ly(dat5, x = ~positive, y = ~negative, size = ~num_qns, color = ~Author, 
	type = 'scatter', mode = 'markers', marker = list(opacity = 0.5)) %>% 
    layout(title = '<b>Sentiments and Num of Questions per Author</b>',
    	   xaxis = list(title = '<b>Positive words</b>', showgrid = FALSE),
    	   yaxis = list(title = '<b>Negative words</b>', showgrid = FALSE),
    	   showlegend = FALSE) %>% 
    add_annotations(
            text = paste(dat5$Author, '\n', dat5$num_qns),
            xref = "x",
            yref = "y",
            showarrow = TRUE,
            arrowsize = 0.5,
            ax = 40,
            ay = -60)

#p

#export(p, file = '/Users/qinqingao/Documents/GitHub/spring2018-project1-ginnyqg/figs/Bubble.png')
```

* With the same method, explore relationship between authors' use of questions in sentences and toward their total volume of texts.

```{r}
#sentence length
spooky$sen_length <- str_length(spooky$text)

dat6 <- mutate(spooky, sen_length = spooky$sen_length)
dat7 <- aggregate(dat6$sen_length, by = list(Author = dat6$author), FUN = sum)

#rename column name
colnames(dat7)[which(names(dat7) == 'x')] <- 'sen_length'

#join tables by author
dat8 <- inner_join(dat7, dat5, by = 'Author')

new_num_texts <- melt(num_texts)
colnames(new_num_texts) <- c('Author', 'num_text')

dat9 <- inner_join(new_num_texts, dat8, by = 'Author')
dat9

#plot bubble chart: Num of Texts, Questions, Sentence Length per Author
p2 <- plot_ly(dat9, x = ~num_text, y = ~num_qns, size = ~sen_length, color = ~Author, 
	type = 'scatter', mode = 'markers', marker = list(opacity = 0.5)) %>% 
    layout(title = '<b>Num of Texts, Questions, Sentence Length per Author</b>',
    	   xaxis = list(title = '<b>Num of Texts</b>', showgrid = FALSE),
    	   yaxis = list(title = '<b>Num of Questions</b>', showgrid = FALSE),
    	   showlegend = FALSE) %>% 
    add_annotations(
            text = paste(dat9$Author, '\n', prettyNum(dat9$sen_length, big.mark = ',', scientific = FALSE)),
            xref = "x",
            yref = "y",
            showarrow = TRUE,
            arrowsize = 0.5,
            ax = 40,
            ay = -60)


#p2

#export(p2, file = '/Users/qinqingao/Documents/GitHub/spring2018-project1-ginnyqg/figs/Bubble_num_text_qns_sent.png')
```







