---
Title: "SPOOKY Data Analysis"
Author: "Ginny Gao"
Date: "Feb 03, 2018"
Output:
  pdf_document: default
  html_document: default
---

# Goal

Goal of this project is to identy authors in `spooky` dataset. Data consists of excerpts of texts written by popular horror authors [Edgar Allan Poe](https://en.wikipedia.org/wiki/Edgar_Allan_Poe), [HP Lovecraft](https://en.wikipedia.org/wiki/H._P._Lovecraft), and [Mary Wollstonecraft Shelley](https://en.wikipedia.org/wiki/Mary_Shelley).


# Load packages and read the data

## Setup the libraries if not already installed


```{r, message = F, warning = F}
packages.used <- c("ggplot2", "plotrix", "waffle", "dplyr", "tibble", "tidyr",  "stringr", "tidytext", "topicmodels", "wordcloud")

# check packages that need to be installed.
packages.needed <- setdiff(packages.used, intersect(installed.packages()[,1], packages.used))

# install additional packages
if(length(packages.needed) > 0) {
  install.packages(packages.needed, dependencies = TRUE, repos = 'http://cran.us.r-project.org')
}


library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
library(stringr)
library(tidytext)
library(topicmodels)
library(wordcloud)
library(plotrix)
library(waffle)
```


## Read in the data
`spooky.csv` in `data` folder, and this Rmd inside `doc` folder.

```{r}
spooky <- read.csv('../data/spooky.csv', as.is = TRUE)
```


# Overview of the dataset

Take a look of first few rows and dimension of the dataset

```{r}
head(spooky)
dim(spooky)
```

How many texts do each author have in the dataset?

```{r}
mytable <- table(spooky$author)
mytable
```

Plot composition of number of texts from 3 authors in pie chart, display counts and percentages


```{r}
lbls <- paste(names(mytable), '\n', mytable, '\n', round(mytable/sum(mytable) * 100, 1), '%', sep = '')
pie3D(mytable, labels = lbls, explode = 0.05, labelcex = 0.8)
```



# Writing Style
## Do some authors use more questions in the texts than others?


* Count number of question marks in texts for spooky  
* Add a field `num_qns` for the counts  
* Wrangle data to show counts for each author  
* Plot a waffle chart to see comparison in use of questions in texts amongst 3 authors.


```{r}
str_count(spooky, '\\?')
dat1 <- mutate(spooky, num_qns = str_count(spooky$text, '\\?'))
dat2 <- data.frame(dat1)
dat3 <- aggregate(dat1$num_qns, by = list(Author = dat1$author), FUN = sum)
dat3

waffle(c('EAP' = dat3[1, 2], 'HPL' = dat3[2, 2], 'MWS' = dat3[3, 2]), rows = 20, size = 0.5, title = 'Count of questions in texts for authors', xlab = '(1 square == 1 question)')
```



# Sentiment analysis
## Positive and negative emotional content comparison in authors' text

Apply sentiment analysis using [bing lexicon](https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html)


```{r}
get_sentiments("bing")

tidy_text <- unnest_tokens(spooky, word, text)
tidy_text_sentiment <- tidy_text %>% inner_join(get_sentiments('bing'))
head(tidy_text_sentiment, 100)


dat4 <- table(tidy_text_sentiment$sentiment, tidy_text_sentiment$author)
dat4


pyramid.plot(dat4[1,c(1:3)], dat4[2,c(1:3)], top.labels = NULL, show.values = TRUE, ndig = 0, main = 'Author by Sentiments', unit = c('Negative', 'Positive'), ppmar = c(4, 4, 4, 4), laxlab = FALSE, raxlab = FALSE)
legend('topright', legend = c("EAP", "HPL", "MWS"), col = c("red", "green", "blue"), lty = 1, bty = 'n', lwd = 8, cex = 0.8, horiz=TRUE)
```








